Exact expressions for double descent and implicit regularization \\ via surrogate random design

### 2022-08-23 design/bayesian_experiments-appendix.tex
Commit original repos

This section presents additional details and experimental results omitted from
the main body of the paper. In addition to the \texttt{mg\_scale} dataset presented in
Section~\ref{s:experiments}, we also benchmarked on three other data sets
described in Table~\ref{tab:libsvm-datasets}.
\begin{table}[ht]
    \centering
    \caption{Datasets used in the experiments \citep{libsvm}.}
    \label{tab:libsvm-datasets}
    \begin{tabular}[t]{lcccc}
        \toprule
        & \texttt{mg\_scale} & \texttt{bodyfat\_scale} & \texttt{mpg\_scale} & \texttt{housing\_scale} \\
        \midrule
        $n$ & 1385 & 252 & 392 & 506 \\
        $d$ & 6 & 14 & 7 & 13\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\textwidth]{../bayesian_figures/obj_grid.pdf}
    \caption{A-optimality values achieved by the methods compared. In all cases
        considered, we found our method (without SDP) to be superior to independent
        sampling methods like uniform and predictive length sampling. After paying the price
        to solve an SDP, our method (with SDP) is able to consistently match the performance
        of a greedy method which has been noted
        \citep{chamon2017approximate} to work well empirically.}
    \label{f:obj-grid}
\end{figure}

@@

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\textwidth]{../bayesian_figures/runtime_grid.pdf}
    \caption{Runtimes of the methods compared. Our method (without SDP) is
        within an order of magnitude of greedy bottom up and faster in 3 out of
        4 cases. The gap between our method with and without SDP is
        attributable to the SDP solver, making investigation of more efficient
        solvers and approximate solutions an interesting direction for future
        work.
    }
    \label{f:runtimes}
\end{figure}

@@

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\textwidth]{../bayesian_figures/ratios_grid.pdf}
    \caption{The ratio controlled by Lemma~\ref{l:guarantees}. This ratio converges
        to $1$ as $k \to n$ and is close to $1$ across all
        real world datasets,
        suggesting that $f_{\A}(\frac{k}{n} \Sigmab_\X)$
        is an appropriate problem-dependent scale for Bayesian A-optimal
        experimental design.
    }
    \label{f:ratios}
\end{figure}

@@

>>>>

\section*{Additional Experiments}

This section presents additional details and experimental results omitted from
the main body of the paper. In addition to the \texttt{mg\_scale} dataset presented in
Section~\ref{s:experiments}, we also benchmarked on three other data sets
described in Table~\ref{tab:libsvm-datasets}.
\begin{table}[ht]
    \centering
    \caption{\cite{libsvm} datasets used in experiments}
    \label{tab:libsvm-datasets}
    \begin{tabular}[t]{lcccc}
        \toprule
        & \texttt{mg\_scale} & \texttt{bodyfat\_scale} & \texttt{mpg\_scale} & \texttt{housing\_scale} \\
        \midrule
        $n$ & 1385 & 252 & 392 & 506 \\
        $d$ & 6 & 14 & 7 & 13\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\textwidth]{bayesian_figures/obj_grid.pdf}
    \caption{A-optimality values achieved by the methods compared. In all cases
        considered, we found our method (without SDP) to be superior to independent
        sampling methods like uniform and predictive length sampling. After paying the price
        to solve an SDP, our method (with SDP) is able to consistently match the performance
        of a greedy method which has been noted
        \cite{chamon2017approximate} to work well empirically.}
    \label{f:obj-grid}
\end{figure}

@@

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\textwidth]{bayesian_figures/runtime_grid.pdf}
    \caption{Runtimes of the methods compared. Our method (without SDP) is
        within an order of magnitude of greedy bottom up and faster in 3 out of
        4 cases. The gap between our method with and without SDP is
        attributable to the SDP solver, making investigation of more efficient
        solvers and approximate solutions an interesting direction for future
        work.
    }
    \label{f:runtimes}
\end{figure}

@@

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\textwidth]{bayesian_figures/ratios_grid.pdf}
    \caption{The ratio controlled by Lemma~\ref{l:guarantees}. This ratio converges
        to $1$ as $k \to n$ and is close to $1$ across all
        real world datasets,
        suggesting that $f_{\A}(\frac{k}{n} \Sigmab_\X)$
        is an appropriate problem-dependent scale for Bayesian A-optimal
        experimental design.
    }
    \label{f:ratios}
\end{figure}

@@

In addition to \texttt{mg\_scale}, the results for the low effective dimension
($d_\lambda$) ridge regression experiments on the remaining \texttt{libsvm}
datasets are shown in Figure~\ref{f:small_deff_ridge_and_a_opt}.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\textwidth]{bayesian_figures/small_deff_ridge_and_a_opt.pdf}
    \caption{Ridge regression in low effective dimension on \texttt{libsvm} datasets.}
    \label{f:small_deff_ridge_and_a_opt}
\end{figure}

### 2022-08-23 design/bayesian_experiments.tex
Commit original repos

\begin{figure*}
    \centering
    \hspace{-0.35cm}
    \includegraphics[width=\textwidth]{Figures/mg_combined.pdf}
     % \includegraphics[width=0.51\textwidth]{bayesian_figures/objectives-mg_scale-full.pdf}\hspace{-2mm}\nobreak\includegraphics[width=0.51\linewidth]{bayesian_figures/ratios-mg_scale.pdf}
    % \includegraphics[width=0.49\textwidth]{bayesian_figures/objectives-mg_scale-zoom.pdf}
    \caption{(left) A-optimality value obtained by the various methods on
        the \texttt{mg\_scale} dataset \citep{libsvm} with
        prior precision $\A = 10^{-5}\, \I$,\quad (right)
A-optimality value for our method (with and without SDP) divided by
$f_{\A}(\frac kn\Sigmab_{\X})$, the baseline estimate suggested by Theorem \ref{t:q1}.}
    \label{f:experiments}
\end{figure*}

\section{Experiments}\label{s:experiments}
We confirm our theoretical
results with experiments on real world data from \texttt{libsvm} datasets
\citep{libsvm} (more details in Appendix~\ref{a:experiments}).
For all our experiments, the prior precision matrix is set to $\A = n^{-1} \I$
and we consider sample sizes $k \in [d, 5d]$. Each experiment is
averaged over 25 trials and bootstrap 95\% confidence intervals are shown.
% We provide open source implementations for efficient RDPP sampling
% and reproducing our results at \Red{TODO: fill in after blind review}.
The quality of our method, as measured by the A-optimality
criterion,
\[f_{\A}(\X_S^\top \X_S) = \tr \left((\X_S^\top \X_S + \A)^{-1}\right),\]
is compared against several baselines and recently proposed methods
for A-optimal design that have been shown to perform well in
practice. Note that none of these algorithms come with theoretical
guarantees as strong as those offered by our approach. The list of
implemented methods is as follows: 
\begin{description}
    \item[Our method (with SDP)] uses the efficient algorithms
        developed in proving Theorem~\ref{t:q2} to sample
        $\DPPreg{p}(\X,\A)$ constrained to subset size $k$
        with $p = w^*$, see \eqref{eq:sdp},
        obtained using a         recently developed first order convex cone solver called Splitting
        Conical Solver \citep[SCS, see][]{o2016conic}.
        We chose SCS because it can handle the SDP constraints in
        \eqref{eq:sdp} and has provable termination guarantees, while
        also finding solutions faster \citep{o2016conic} than alternative
        off-the-shelf optimization software libraries such as SDPT3
        and Sedumi.

@@

@@

Consistent with prior observations
\citep{tractable-experimental-design,chamon2017approximate}, the greedy bottom up
method achieves surprisingly good performance, despite the limited
theoretical guarantees it offers. However, if our method is used
in conjunction with an SDP solution, then we are able to match and
even slightly exceed the performance of the greedy bottom up
method. Furthermore, the overall run-time costs (see Appendix~\ref{a:experiments})
between the two are comparable. As the majority of the runtime of our
method (with SDP) is occupied by solving the SDP, an interesting future direction
is to investigate alternative solvers such as interior point methods as well
as terminating the solvers early once an approximate solution is reached.

@@

>>>>

\begin{figure*}
    \centering
    \hspace{-0.35cm}
    \includegraphics[width=\textwidth]{aistats/Figures/mg_combined.pdf}
     % \includegraphics[width=0.51\textwidth]{bayesian_figures/objectives-mg_scale-full.pdf}\hspace{-2mm}\nobreak\includegraphics[width=0.51\linewidth]{bayesian_figures/ratios-mg_scale.pdf}
    % \includegraphics[width=0.49\textwidth]{bayesian_figures/objectives-mg_scale-zoom.pdf}
    \caption{(left) A-optimality value obtained by the various methods on
        the \texttt{mg\_scale} dataset \cite{libsvm} with
        prior precision $\A = 10^{-5}\, \I$,\quad (right)
A-optimality value for our method (with and without SDP) divided by
$f_{\A}(\frac kn\Sigmab_{\X})$, the baseline estimate suggested by Theorem \ref{t:q1}.}
    \label{f:experiments}
\end{figure*}

\section{Experiments}\label{s:experiments}

In this section we demonstrate that our method is capable of finding better
experimental designs $S \subset [n]$ on data matrices $\X$ occuring in on real
world data from \texttt{libsvm} datasets \citep{libsvm} (more details in
Appendix~\ref{a:experiments}).
We compare our method against the following baselines and recently proposed methods
for A-optimal design that have been shown to perform well in
practice:
\begin{description}
    \item[Our method (with SDP)] uses the efficient algorithms
        developed in proving Theorem~\ref{t:q2} to sample
        $\DPPreg{p}(\X,\A)$ constrained to subset size $k$
        with $p = w^*$, see \eqref{eq:sdp},
        obtained using a         recently developed first order convex cone solver called Splitting
        Conical Solver (SCS) \citep{o2016conic}.
        We chose SCS because it can handle the SDP constraints in
        \eqref{eq:sdp} and has provable termination guarantees, while
        also finding solutions faster \citep{o2016conic} than alternative
        off-the-shelf optimization software libraries such as SDPT3 and Sedumi.

@@

\subsection{Bayesian A-optimal design}

Consistent with prior work \citep{mariet2017elementary,chaloner1984optimal}
in optimal Bayesian experimental design,
we consider the task of A-optimal design:
given a data matrix $\X \in \R^{n \times d}$, select
$S \subset [n]$ of size $\lvert S \rvert = k$ such that
the A-optimality
criterion,
\[f_{\A}(\X_S^\top \X_S) = \tr \left((\X_S^\top \X_S + \A)^{-1}\right),\]
is small.
All our experiments set the prior precision matrix to $\A = n^{-1} \I$
and we consider sample sizes $k \in [d, 5d]$.
Each experiment is averaged over 25 trials and bootstrap 95\% confidence
intervals are shown.

@@

Consistent with prior observations
\citep{chamon2017approximate,tractable-experimental-design}, the greedy bottom up
method achieves surprisingly good performance, despite the limited
theoretical guarantees it offers. However, if our method is used
in conjunction with an SDP solution, then we are able to match and
even slightly exceed the performance of the greedy bottom up
method. Furthermore, the overall run-time costs (see Appendix~\ref{a:experiments})
between the two are comparable. As the majority of the runtime of our
method (with SDP) is occupied by solving the SDP, an interesting future direction
is to investigate alternative solvers such as interior point methods as well
as terminating the solvers early once an approximate solution is reached.

@@

\subsection{Subsampled ridge regression}

\begin{figure}[htbp]
    \centering
    % \hspace{-0.85cm}
    \includegraphics[width=0.8\textwidth]{bayesian_figures/mg_scale_ridge.pdf}
    \caption{Mean-squared error for subsampled ridge estimates
        on \texttt{mg\_scale}
    }
    \label{fig:low_d_eff}
\end{figure}

As experimental design is oftentimes employed as a data reduction step within a
predictive modelling pipeline, we validate the practical utility of our
algorithm by demonstrating superior performance for row subset selection in a
ridge regression task on $\texttt{mg\_scale}$.
We chose $\A = \lambda \I$ such that $\approx 95\%$ of the spectral mass was
contained in eigenvalues exceeding $\lambda$, resulting in an effective
dimension $d_\A \approx 3.97 \ll 6 = d$.  Because $d_\A$ is smaller than $d$,
the new theoretical results in this paper (Theroem~\ref{t:q1}) provide tighter
guarantees than previous bounds which depended only on $d$.

We first use the various methods to produce a design $S \subset [n]$,
form the subsampled ridge estimator $\hat{\w}_S = (\X_S^\top \X_S + \A)^{-1}
\X_S^\top \y_S$ (which is also the posterior mean of $\w$ given $\y_S$),
and plots its $\ell_2$ distance from the full unsubsampled
estimator $\hat{\w}_{[n]}$ (i.e. the posterior mean of $\w$ given $\y_{[n]}$)
in Figure~\ref{fig:low_d_eff} ($100$ trials with $95\%$ confidence intervals).
Our method outperforms all of the other methods compared, achieveing a lower
relative approximation error as the sample size increases.

### 2022-08-23 design/continuous_colt/continuous-empirical.tex
Commit original repos

\subsection{Empirical analysis of asymptotic consistency}
\label{s:empirical}

Theorem \ref{t:asymptotic} states that our surrogate expressions for the MSE under certain Gaussian designs are asymptotically consistent with the multiplicative error rate of $O(1/d)$. 
In this section, we show strong empirical evidence that this fact extends to
the setting not covered by the theorem: under-determined regime (i.e.,
$n<d$) with a non-isotropic Gaussian distribution (i.e., $\Sigmab\neq
\I$). We break down our analysis into verifying two
conjectures which are of independence interest to multivariate
Gaussian analysis. The first conjecture addresses the variance term in
the MSE and postulates an asymptotically
consistent formula for the expected Moore-Penrose pseudo-inverse
of the singular Wishart distribution. Recall that matrix
$\W\sim\Wc(\Sigmab,n)$ is distributed according to the Wishart
distribution with $n$ degrees of freedom if it can be decomposed as
$\W=\X^\top\X$, where $\X\sim\Nc(\zero,\Sigmab)^n$.
\begin{conjecture}[Pseudo-inverse of singular Wishart]\label{c:wishart}
  Fix $n/d<1$ and let $\W\sim\Wc(\Sigmab,n )$, where
$\Sigmab$ is $d\times d$ positive definite with condition number bounded by a constant.
  Then:
\begin{align}
\bigg|\frac{\E\big[\tr(\W^\dagger)\big]}{\Vc(\Sigmab,n)} -1\bigg|=
  O(1/d)\qquad\text{for}\quad\Vc(\Sigmab,n)=\tr\big((\Sigmab+\lambda_n\I)^{-1}\big)\frac{1-\alpha_n}{d-n},\label{eq:wishart}
\end{align}
where $\lambda_n\geq 0$ satisfies $n=\tr(\Sigmab(\Sigmab+\lambda_n\I)^{-1})$ and
$\alpha_n=\det(\Sigmab(\Sigmab+\lambda_n\I)^{-1})$.
\end{conjecture}
Our second conjecture involves the projection onto the
orthogonal complement of a Gaussian sample $\X$, i.e., the matrix
$\I-\X^\dagger\X$, and addresses the bias term in the MSE.
\begin{conjecture}[Gaussian orthogonal projection]\label{c:projection}
  Fix $n/d<1$ and let $\X\sim \Nc(\zero,\Sigmab)^n$, where
$\Sigmab$ is $d\times d$ positive definite with condition number bounded by a constant. Then:
\begin{align}
\sup_{\w\in\R^d\backslash\{\zero\}}\bigg|\frac{\w^\top\E[\I-\X^\dagger\X]\w}{\w^\top
  \Bc(\Sigmab,n)\w} - 1\bigg| = O(1/d)\qquad\text{for}\quad
  \Bc(\Sigmab,n) =
  (\Sigmab+\lambda_n\I)^{-1}\frac{d-n}{\tr((\Sigmab+\lambda_n\I)^{-1})},\label{eq:projection}  
\end{align}
where $\lambda_n\geq 0$ satisfies $n=\tr(\Sigmab(\Sigmab+\lambda_n\I)^{-1})$.
\end{conjecture}
Note that the surrogate expression for the mean squared
error can be written as:
\[
\Mc(\Sigmab_\mu,\w^*,\sigma^2,n)=\sigma^2 \Vc(\Sigmab_\mu,n)
+ \w^{*\top}\Bc(\Sigmab_\mu,n)\w^*.
\]
So, if the conjectures are true,
this would immediately imply that the asymptotic consistency claim
given in Theorem \ref{t:asymptotic} for
the surrogate Gaussian MSE extends to
the under-determined setting with arbitrary covariance $\Sigmab$.

Furthermore, both of these conjectures are related to open problems
which have been extensively studied in the literature. 
With respect to Conjecture~\ref{c:wishart}, \cite{srivastava2003} first derived the probability
density function of a singular Wishart
distribution, and \cite{cook2011} computed the first and second moments of
generalized inverses of a singular Wishart distribution. However, for the
Moore-Penrose pseudo-inverse and arbitrary covariance $\Sigmab$,
\cite{cook2011} claims that the quantities required
to express the mean ``do not have tractable closed-form representation.''
%
Conjecture~\ref{c:projection} has connections to directional statistics.
Using the SVD, we have the equivalent representation $\X^\dagger \X = \V \V^\top$
where $\V$ is an element of the Stiefel manifold $V_{n,d}$ (i.e., orthonormal
$n$-frames in $\R^d$).
The distribution of $\V$ is known as the matrix angular central
Gaussian (MACG) distribution \cite{chikuse1990matrix}. While prior work
has considered high dimensional limit theorems \cite{CHIKUSE1991145}
as well as density estimation and hypothesis testing \cite{CHIKUSE1998188}
on $V_{n,d}$, they only analyzed the invariant measure
(which corresponds in our setting to $\Sigmab = \I$),
and to our knowledge a closed form expression of $\E[\V\V^\top]$ where
$\V$ is distributed according to MACG with
arbitrary $\Sigmab$ remains an open question.

For verifying these two conjectures, it suffices to only consider
diagonal covariance matrices $\Sigmab$.  This is because if $\Sigmab = \Q \D
\Q^\top$ is its eigendecomposition and $\X\sim\Nc(\zero, \Q\D\Q^\top)^n$, then
we have for $\W \sim \Wc(\Sigmab, n)$ that $\W
\overset{d}{=} \X^\top \X$ and hence, defining
$\Xt\sim\Nc(\zero,\D)^n$, by linearity and unitary invariance of trace,
\begin{align*}
  \E[\tr(\W^\dagger)]
  &= \tr\big( \E[(\X^\top\X)^\dagger] \big)
  = \tr\Big( \Q\E\big[(\Xt^\top\Xt)^\dagger\big]\Q^\top \Big)
  = \tr\Big( \E\big[(\Xt^\top\Xt)^\dagger\big] \Big)
  = \E\left[\tr \big((\Xt^\top\Xt)^\dagger\big) \right].
\end{align*}
Similarly, we have that
$\E[\X^\dagger\X]=\Q\E\big[\Xt^\dagger\Xt\big]\Q^\top$, and a simple
calculation shows that the expression in Conjecture~\ref{c:projection}
is also independent of the choice of matrix $\Q$.
Thus, we empirically validate our conjectures for diagonal matrices $\Sigmab$
with several different eigenvalue decay profiles. Denoting
$\lambda_1,...,\lambda_d$ as the eigenvalues of $\Sigmab$, we consider
the following decays:
\begin{itemize}
  \item \texttt{diag\_linear}: linear decay, $\lambda_{i} = b-a i$;
  \item \texttt{diag\_exp}: exponential decay, $\lambda_{i} = b\,10^{-
      a i} $;
  \item \texttt{diag\_poly}: fixed-degree polynomial decay, $\lambda_{i} = (b-a i)^2$;
  \item \texttt{diag\_poly\_2}: variable-degree polynomial decay, $\lambda_i = b i^{-a}$.
\end{itemize}
The constants $a$ and $b$ are chosen to ensure $\lambda_{\text{max}}(\Sigmab) = 1$ and
$\lambda_{\text{min}}(\Sigmab) = 10^{-4}$ (i.e., the condition number
$\kappa(\Sigmab) = 10^{4}$ remains constant).
Figure~\ref{fig:eig-decays} illustrates an example of these decay profiles
for $d=100$.

\begin{figure}[H]
    \includegraphics[width=\textwidth]{continuous_figures/decays.pdf}
  \caption{Scree-plots of $\Sigmab$ for the eigenvalue decays examined
    in our empirical valuations. Here $d=100$ for visualization, whereas
    our experiments increase $d$ while preserving the ratio $n/d$ and
    the decay profile,
    with $\lambda_{\text{max}}(\Sigmab) = 1$ to
    $\lambda_{\text{min}}(\Sigmab) = 10^{-4}$.}
  \label{fig:eig-decays}
\end{figure}

We verify our conjectures by incrasing $d$ while keeping the aspect ratio
$n/d$ fixed and examining the rate of decay of the quantities asserted
in the conjectures. As no closed form expressions are available for
the expectations in the conjectures, we estimate $\E\big[\tr(\W^\dagger)\big]$ (for
Conjecture~\ref{c:wishart}) and $\E[\I-\X^\dagger\X]$ (for
Conjecture~\ref{c:projection}) through Monte Carlo sampling. To ensure that
estimation noise is sufficiently small, we continually increase the number
of Monte Carlo samples until the bootstrap confidence intervals are
within $\pm 12.5\%$ of the quantities in \eqref{eq:wishart} and
\eqref{eq:projection}. We found that while
Conjecture~\ref{c:wishart} required a relatively small number of
trials (up to one thousand),
estimation noise was much larger in Conjecture~\ref{c:projection} and
necessitated over two million trials to obtain good estimates
near $d=100$.

% For Conjecture~\ref{c:wishart}, sampling $\W\sim\Wc(\Sigmab,n)$ is
% elementary and the only nontrivial aspect of computing $\Vc(\Sigmab, n)$
% is solving for $\lambda_n$ such that
% $n = \tr(\Sigmab(\Sigmab + \lambda_n \I)^{-1})$. But notice that
% if we let $\tau_i > 0$ be the eigenvalues of $\Sigmab \succ 0$,
% then
% \begin{align*}
%   \tr(\Sigmab(\Sigmab + \lambda \I)^{-1})
%   &= \sum_{i=1}^d \frac{\tau_i}{\tau_i + \lambda}
% \end{align*}
% which is equal to $d > n$ when $\lambda = 0$ and monotonically decreasing in
% $\lambda$. Hence, $\lambda_n$ is unique and can be found using bisection.

\begin{figure}[h]
    \includegraphics[width=\textwidth]{continuous_figures/wishart_var.pdf}
  \caption{
    Empirical verification of Conjecture~\ref{c:wishart}. We show
    the quantity $\big|\E[\tr(\W^\dagger)]\,\Vc(\Sigmab,n)^{-1} -1\big|$
    as $d$ increases for various aspect ratios $n/d$. Consistent with our conjecture,
    a $O(1/d)$ decay (linear with slope $-1$ on a log-log plot) is
    exhibited across all eigenvalue decay profiles and
    aspect ratios investigated.
  }
  \label{f:conj-wishart}
\end{figure}

The results of empirically validating Conjecture~\ref{c:wishart} are
illustrated in Figure~\ref{f:conj-wishart}, where we performed
Monte Carlo estimation of $\E\big[\tr(\W^\dagger)\big]$ and
plot $\big|\E[\tr(\W^\dagger)]\,\Vc(\Sigmab,n)^{-1} -1\big|$
as $d$ increases from $10$ to $1000$, across a range of aspect ratios $n/d$ and eigenvalue
decay profiles for $\Sigmab$. Confidence intervals are estimated by
bootstrapping. We observe that on log-log axes all of the plots are decreasing with
a linear $-1$ slope, consistent with the $O(1/d)$ rate predicted
by Conjecture~\ref{c:wishart}.

Conjecture \ref{c:projection} is handled similarly, by sampling
$\X\sim\mu^n$ where $\mu=\Nc(\zero,\Sigmab )$ to obtain a Monte Carlo estimate of
$\E[\I-\X^\dagger\X]$. To handle the supremum over $\w$, notice that
we can rewrite \eqref{eq:projection} as a spectral norm
\begin{align}
  \sup_{\w\in\R^d\backslash\{\zero\}}\bigg|\frac{\w^\top\E[\I-\X^\dagger\X]\w}{\w^\top
  \Bc(\Sigmab,n)\w} - 1\bigg|
\  =\ \big\|\E[\I-\X^\dagger\X]\Bc(\Sigmab,n)^{-1}
  - \I\big\|.
  \label{eq:proj-to-operator-norm}
\end{align}
Confidence intervals can now be constructed
using existing methods for constructing operator norm
confidence intervals, and our results use the bootstrapping
method described in \cite{lopes2019bootstrapping}.

Figure~\ref{f:conj-bias} shows
how $\big\|\E[\I-\X^\dagger\X]\Bc(\Sigmab,n)^{-1} - \I\big\|$ decays as we
hold the aspect ratio $n/d$ fixed and increase $d$ between $10$ and
$100$ across the listed
eigenvalue decay profiles and aspect ratios. Again, we observe on
log-log axes a linear decay with slope $-1$ consistent with the $O(1/d)$
rate posed by Conjecture~\ref{c:projection}. Note that the range of
$d$ is smaller than in Figure \ref{f:conj-wishart} because the large
number of Monte Carlo samples (up to two million) required for this
experiment made the computations much more expensive.

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=\textwidth]{continuous_figures/proj_bias.pdf}
  \end{center}
  \caption{
    Empirical results validating Conjecture~\ref{c:projection}.
We show how
    $\big\|\E[\I-\X^\dagger\X]\Bc(\Sigmab,n)^{-1} - \I\big\|$
    (which by Equation~\ref{eq:proj-to-operator-norm}
    is equal to the quantity controlled by Conjecture~\ref{c:projection})
    decays as $d$ increases and observe a linear slope consistent with the conjectured $O(1/d)$ rate.
  }
  \label{f:conj-bias}
\end{figure}

>>>>

\begin{figure}[t]
  \includegraphics[width=\textwidth]{continuous_figures/decays.pdf}
  \vspace{-1cm}
  \caption{Scree-plots of $\Sigmab$ for the eigenvalue decays examined
    in our empirical valuations. Here $d=100$ for visualization, whereas
    our experiments increase $d$ while preserving the ratio $n/d$ and
    the decay profile,
    with $\lambda_{\text{max}}(\Sigmab) = 1$ to
    $\lambda_{\text{min}}(\Sigmab) = 10^{-4}$.}
  \label{fig:eig-decays}
\end{figure}

% Denoting
% $\lambda_1,...,\lambda_d$ as the eigenvalues of $\Sigmab$, w
Letting $\lambda_i(\Sigmab)$ be the $i$th largest eigenvalue of
$\Sigmab$, we consider the
following eigenvalue profiles:
\begin{itemize}
  \item \texttt{diag\_linear}: linear decay, $\lambda_i(\Sigmab)= b-a i$;
  \item \texttt{diag\_exp}: exponential decay, $\lambda_i(\Sigmab) = b\,10^{- a i} $;
  \item \texttt{diag\_poly}: fixed-degree polynomial decay, $\lambda_i(\Sigmab) = (b-a i)^2$;
  \item \texttt{diag\_poly\_2}: variable-degree polynomial decay, $\lambda_i(\Sigmab) = b i^{-a}$.
\end{itemize}
The constants $a$ and $b$ are chosen to ensure $\lambda_{\text{max}}(\Sigmab) = 1$ and
$\lambda_{\text{min}}(\Sigmab) = 10^{-4}$ (i.e., the condition number
$\kappa(\Sigmab) = 10^{4}$ remains constant).

We verify our conjectures by increasing $d$ while keeping the aspect ratio
$n/d$ fixed and examining the rate of decay of the quantities asserted
in the conjectures. As no closed form expressions are available for
the expectations in the conjectures, we estimate $\E\big[\tr(\W^\dagger)\big]$ (for
Conjecture~\ref{c:wishart}) and $\E[\I-\X^\dagger\X]$ (for
Conjecture~\ref{c:projection}) through Monte Carlo sampling. To ensure that
estimation noise is sufficiently small, we continually increase the number
of Monte Carlo samples until the bootstrap confidence intervals are
within $\pm 12.5\%$ of the quantities in \eqref{eq:wishart} and
\eqref{eq:projection}. We found that while
Conjecture~\ref{c:wishart} required a relatively small number of
trials (up to one thousand),
estimation noise was much larger in Conjecture~\ref{c:projection} and
necessitated over two million trials to obtain good estimates
near $d=100$.

% For Conjecture~\ref{c:wishart}, sampling $\W\sim\Pc\Wc(\Sigmab,n)$ is
% elementary and the only nontrivial aspect of computing $\Vc(\Sigmab, n)$
% is solving for $\lambda_n$ such that
% $n = \tr(\Sigmab(\Sigmab + \lambda_n \I)^{-1})$. But notice that
% if we let $\tau_i > 0$ be the eigenvalues of $\Sigmab \succ 0$,
% then
% \begin{align*}
%   \tr(\Sigmab(\Sigmab + \lambda \I)^{-1})
%   &= \sum_{i=1}^d \frac{\tau_i}{\tau_i + \lambda}
% \end{align*}
% which is equal to $d > n$ when $\lambda = 0$ and monotonically decreasing in
% $\lambda$. Hence, $\lambda_n$ is unique and can be found using bisection.

\begin{figure}[ht]
  \includegraphics[width=\textwidth]{continuous_figures/wishart_var.pdf}
  \includegraphics[width=\textwidth]{continuous_figures/proj_bias.pdf}
  \vspace{-1cm}
  \caption{
    Empirical verification of Conjectures~\ref{c:wishart} (top) and
    \ref{c:projection} (bottom).
    We show the quantities
    $|\E[\tr(\W^\dagger)]\,\Vc(\Sigmab,n)^{-1} -1|$ and
    $\|\E[\I-\X^\dagger\X]\Bc(\Sigmab,n)^{-1} - \I\|$, respectively,
    as $d$ increases for various aspect ratios $n/d$.
    Consistent with our conjectures,
    a $O(1/d)$ decay (linear with slope $-1$ on a log-log plot) is
    exhibited across all eigenvalue decay profiles and
    aspect ratios investigated.
  }
  \label{f:conj-wishart}
\end{figure}

The results of empirically validating Conjecture~\ref{c:wishart} are
illustrated in Figure~\ref{f:conj-wishart} (top), where we performed
Monte Carlo estimation of $\E\big[\tr(\W^\dagger)\big]$ and
plot $\big|\E[\tr(\W^\dagger)]\,\Vc(\Sigmab,n)^{-1} -1\big|$
as $d$ increases from $10$ to $1000$, across a range of aspect ratios $n/d$ and eigenvalue
decay profiles for $\Sigmab$. Confidence intervals are estimated by
bootstrapping. We observe that on log-log axes all of the plots are decreasing with
a linear $-1$ slope, consistent with the $O(1/d)$ rate predicted
by Conjecture~\ref{c:wishart}.

Conjecture \ref{c:projection} is handled similarly, by sampling
$\X\sim\mu^n$ where $\mu=\Nc(\zero,\Sigmab )$ to obtain a Monte Carlo estimate of
$\E[\I-\X^\dagger\X]$. To handle the supremum over $\w$, notice that
we can rewrite \eqref{eq:projection} as a spectral norm
\begin{align}
  \sup_{\w\in\R^d\backslash\{\zero\}}\bigg|\frac{\w^\top\E[\I-\X^\dagger\X]\w}{\w^\top
  \Bc(\Sigmab,n)\w} - 1\bigg|
\  =\ \big\|\E[\I-\X^\dagger\X]\Bc(\Sigmab,n)^{-1}
  - \I\big\|.
  \label{eq:proj-to-operator-norm}
\end{align}
Confidence intervals can now be constructed
using existing methods for constructing operator norm
confidence intervals, and our results use the bootstrapping
method described by \cite{lopes2019bootstrapping}.

Figure~\ref{f:conj-wishart} (bottom) shows
how $\big\|\E[\I-\X^\dagger\X]\Bc(\Sigmab,n)^{-1} - \I\big\|$ decays as we
hold the aspect ratio $n/d$ fixed and increase $d$ between $10$ and
$100$ across the listed
eigenvalue decay profiles and aspect ratios. Again, we observe on
log-log axes a linear decay with slope $-1$ consistent with the $O(1/d)$
rate posed by Conjecture~\ref{c:projection}. Note that the range of
$d$ is smaller than in Figure \ref{f:conj-wishart} (top) because the large
number of Monte Carlo samples (up to two million) required for this
experiment made the computations much more expensive.
